---
title: "Food Factors"
output: html_document
---

# Load and Format the Data

```{r}
library(MASS)
library(plyr)
library(corrplot)
library(RColorBrewer)
suppressWarnings(library(ggplot2))
library(tidyverse)
```

```{r}
dat <- read.csv("z_trial_data.csv")[,-1]
dat <- as_tibble(dat)
```

To make the visuals more interpretable, we will flip 4 of the measures. This is to make these four measures become positively related to the other measures.

* Sugar => Low-Sugar
* Disgusting => Not Disgusting
* Healthy => Unhealthy
* Vitamins => Low vitamins

```{r}
rdat <- dat %>% select(stimulus, choice.rating, starts_with("rating"))

# Rename
colnames(rdat) <- sub("rating[.]", "", colnames(rdat))
colnames(rdat) <- sub("[.]rating", "", colnames(rdat))

# Flip
rdat$sugar <- 10 - rdat$sugar
rdat$disgusting <- 10 - rdat$disgusting
rdat$healthy <- 10 - rdat$healthy
rdat$vitamins <- 10 - rdat$vitamins
## rename
cinds <- sapply(c("sugar", "disgusting", "healthy", "vitamins"), function(x) which(colnames(rdat) == x))
colnames(rdat)[cinds] <- c("low-sugar", "not-disgusting", "unhealthy", "low-vitamins")

# For use with analysis
rmat <- as.matrix(rdat[,-1])
rownames(rmat) <- rdat$stimulus
rmat.nochoice <- rmat[,-1]

# Shows the ratings/choice for each of the food items
head(rmat)
```


# Correlation Matrix

Get the average of each of the ratings/choice across participants. Then compute and visualize the correlations between each pair of ratings/choice.

```{r}
ave.mat <- rdat %>% 
  group_by(stimulus) %>% 
  summarise_all(mean) %>% 
  select(-stimulus)

M <- cor(ave.mat)
diag(M) <- 0

corrplot(M, method="circle", diag = F, order="hclust", 
         #is.corr=F, cl.lim=c(-0.85,0.85), 
         col=rev(colorRampPalette(brewer.pal(n=11, name="RdBu"))(256)))
```

Manually order based on what we might have expected to see

```{r}
new.ord <- c("familiar", "calories", "fat", "protein", "carbohydrates", "gluten", "low-vitamins", "sodium", "low-sugar", "sweetsaltysavory", "filling", "texture", "not-disgusting", "feel", "othertasty", "tasty", "unhealthy", "choice")

ave.mat2 <- ave.mat[,new.ord]
M <- cor(ave.mat2)
diag(M) <- 0

corrplot(M, method="circle", diag = F, #order="hclust", 
         #is.corr=F, cl.lim=c(-0.85,0.85), 
         col=rev(colorRampPalette(brewer.pal(n=11, name="RdBu"))(256)))
```


# Factor Analysis

We can see that there is a clear structure in the data above. Here, we try to extract that structure by doing a factor analysis.

For the factor analysis, I will remove the choice data to make it only about the ratings.

Please note in the section below, I go through a few different possibilities for the number of factors. I select 3 since that is the best with the CNG method and also because that number best explains choice behavior.

Determine the optimal number of factors (it's 3)

```{r}
nFactors::nCng(as.data.frame(rmat.nochoice), model="factors") 
```

```{r}
cols2 <- colorRampPalette(c("#67001F", "#B2182B", "#D6604D", 
                            "#F4A582", "#FDDBC7", "#FFFFFF", "#D1E5F0", "#92C5DE", 
                            "#4393C3", "#2166AC", "#053061"))(200)
```

Plot 3 factors and there factor scores. I use hierarchical clustering to order the different ratings and make the plot more interpretable.

```{r}
fit <- factanal(rmat.nochoice, 3, scores="regression")

d <- dist(fit$loadings)
hc <- hclust(d, method="average")
dend <- as.dendrogram(hc)
loads <- fit$loadings[order.dendrogram(dend),]
loads <- rbind(choice=loads[rownames(loads) == "choice",], loads[rownames(loads) != "choice",])

# Make tasty first, then health, and finally sweet/savory
loads2 <- loads[,c(2,1,3)]
colnames(loads2) <- sprintf("Factor-%i", 1:3)

# Put the sweet/savory to the end
#loads2 <- rbind(loads2[-c(1:4),], loads2[4:1,])
loads2 <- rbind(loads2[-c(1:4),], loads2[1:4,])

# Just for reporting purposes
sort(loads2[,1], decreasing = T)
sort(loads2[,2], decreasing = T)
sort(loads2[,3], decreasing = T)

# Plot
corrplot::corrplot(loads2, diag=T, col=rev(cols2), cl.pos="n")
```

Do a different arrangement for the audience.

```{r}
rlabs <- c("tasty", "othertasty", "feel", "texture", "not-disgusting", "familiar", 
  "calories", "unhealthy", "fat", "carbohydrates", "low-vitamins", "gluten", "sodium", 
  "protein", "sweetsaltysavory", "filling", "low-sugar")

inds <- sapply(rlabs, function(x) which(rownames(loads2)==x))
corrplot::corrplot(loads2[inds,], diag=T, col=rev(cols2), cl.pos="n")
```
Gather the scores and re-label the columns accordingly.

```{r}
# scores and re-label columns
scores3 <- fit$scores
ind1 <- which.max(fit$loadings[rownames(fit$loadings) == "unhealthy",])
ind2 <- which.max(fit$loadings[rownames(fit$loadings) == "tasty",])
ind3 <- which.max(fit$loadings[rownames(fit$loadings) == "sweetsaltysavory",])
colnames(scores3) <- c("Food.UnHealth", "Food.Taste", "Food.SweetProtein")[c(ind1,ind2,ind3)]
```

Save the data out

```{r}
head(scores3)
```

```{r}
outdf <- data.frame(subjectId=dat$subjectId, stimulus=dat$stimulus, scores3)
write.csv(outdf, file="z_foodrating_factors.csv")
```

